---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

setwd("C:/Users/SAURAV/Downloads/repdata_data_activity")
activity<-read.csv("activity.csv")
```
Loading this data
# Loading and preprocessing the data
```{r}
dim(activity)
names(activity)
head(activity)
str(activity)
#total number of missing data
sum(is.na(activity$steps))/dim(activity)[[1]]
#transforming the date column into date format using lubridate
library(lubridate)
activity$date<-ymd(activity$date)
length(unique(activity$date))
  filename <- "activity.csv"
  
  if(!file.exists(filename))
    {
      download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",destfile="activity.zip")
      unzip(filename)
      activity <- read.csv(filename)
    }
    
  activity <- read.csv(filename)
  str(activity)
  head(activity)
  
```
# What is mean total number of steps taken per day?

## Histogram of the total number of steps taken each day
```{r}
library(dplyr)
library(ggplot2)
act <- aggregate(steps ~ date, data=activity, sum, na.rm=TRUE)
head(act)
activity$date <- as.Date(activity$date, "%Y-%m-%d")
qplot(steps,data=act,color="Red",xlab = "Number of steps", ylab="Interval", main="Total number of steps per day")
```

## What is mean total number of steps taken per day?
```{r, echo = TRUE}
activity_total_steps <- with(activity, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
names(activity_total_steps) <- c("date", "steps")
hist(activity_total_steps$steps, main = "Total number of steps taken per day", xlab = "Total steps taken per day", col = "blue", ylim = c(0,20), breaks = seq(0,25000, by=2500))
mean(activity_total_steps$steps)
median(activity_total_steps$steps)
## Mean and median total number of steps taken per day
```{r}
meansteps<-mean(act$steps)
```
```{r}
mediansteps<-median(act$steps)
```

## What is the average daily activity pattern?

# Time series plot of the average number of steps taken

```{r}
average_daily_activity <- aggregate(activity$steps, by=list(activity$interval), FUN=mean, na.rm=TRUE)
names(average_daily_activity) <- c("interval", "mean")
plot(average_daily_activity$interval, average_daily_activity$mean, type = "l", col="darkblue", lwd = 2, xlab="Interval", ylab="Average number of steps", main="Average number of steps per intervals")
average_daily_activity[which.max(average_daily_activity$mean), ]$interval
stepinterval <- aggregate(steps~interval, data=activity, mean, na.rm=TRUE)
head(stepinterval)
qplot(x = interval,
      y = steps, 
      data = stepinterval,
      geom = "line",
      colour = I("red"), 
      xlab = "5 mins Interval",
      ylab = "Steps taken",
      main = "Time series plot of average number of steps")
```
# Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
maxsteps <- stepinterval$interval[which.max(stepinterval$steps)]
maxsteps
```

## Imputing missing values
sum(is.na(activity$steps))
imputed_steps <- average_daily_activity$mean[match(activity$interval, average_daily_activity$interval)]
activity_imputed <- transform(activity, steps = ifelse(is.na(activity$steps), yes = imputed_steps, no = activity$steps))
total_steps_imputed <- aggregate(steps ~ date, activity_imputed, sum)
names(total_steps_imputed) <- c("date", "daily_steps")
hist(total_steps_imputed$daily_steps, col = "darkblue", xlab = "Total steps per day", ylim = c(0,30), main = "Total number of steps taken each day", breaks = seq(0,25000,by=2500))
mean(total_steps_imputed$daily_steps)
median(total_steps_imputed$daily_steps)

# Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
nas <- sum(is.na(activity$steps))
nas
```
# Strategy for filling in missing values.

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
activity2 <- activity
isna <- is.na(activity2$steps) 
avg_interval <- tapply(activity2$steps, activity2$interval, mean, na.rm=TRUE, simplify = TRUE)
activity2$steps[isna] <- avg_interval[as.character(activity2$interval[isna])]
sum(is.na(activity2))
act2 <- aggregate(steps ~ date, data = activity2, FUN = sum, na.rm = TRUE)
par(mfrow=c(1,2))
## Histogram without the NA values
hist(act2$steps, 
    main = "Total Steps per Day (no-NA)", 
    xlab = "Number of Steps per Day", 
    ylab = "Interval",
    col="green",
    breaks=50)
##Histogram with the orginal dataset
hist(act$steps, 
    main="Total Steps per Day (Original)", 
    xlab="Number of Steps per Day", 
    ylab = "Interval",
    col="orange",
    breaks=50)
```
## Are there differences in activity patterns between weekdays and weekends?
activity$date <- as.Date(strptime(activity$date, format="%Y-%m-%d"))
activity$datetype <- sapply(activity$date, function(x) {
        if (weekdays(x) == "Sábado" | weekdays(x) =="Domingo") 
                {y <- "Weekend"} else 
                {y <- "Weekday"}
                y
        })
        activity_by_date <- aggregate(steps~interval + datetype, activity, mean, na.rm = TRUE)
plot<- ggplot(activity_by_date, aes(x = interval , y = steps, color = datetype)) +

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
par(mfrow=c(1,1))
activity2<- activity2%>%
        mutate(typeofday= ifelse(weekdays(activity2$date)=="Saturday" | weekdays(activity2$date)=="Sunday", "Weekend", "Weekday"))
        
head(activity2)
      
```
Make a panel plot containing a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
fivemin2<- aggregate(steps ~ interval, data = activity2, FUN = mean, na.rm = TRUE)
head(fivemin2)
```
```{r}
ggplot(activity2, aes(x =interval , y=steps, color=typeofday)) +
       geom_line() +
       labs(title = "Average daily steps by type of date", x = "Interval", y = "Average number of steps") +
       facet_wrap(~datetype, ncol = 1, nrow=2)
print(plot)
       labs(title = "Ave Daily Steps (type of day)", x = "Interval", y = "Total Number of Steps") +
       facet_wrap(~ typeofday, ncol = 1, nrow=2)
```
